Last glacial cycle glacier erosion potential in the Alps

### 2021-01-25 articles/erosion/alps-erosion.tex
Add Ian (hoping he accepts), rework the abstract.

\title{Last glacial cycle glacier erosion potential in the Alps}
\author{Julien Seguinot}
\affil{Anafi, Greece}

@@

    The glacial landscape of the Alps has fascinated generations of explorers,
    artists, mountaineers and scientists with its diversity, including
    erosional features of all scales from high-mountain cirques, to steep
    glacial valleys and large overdeepenings. Using previous glacier modelling
    results, and modern observations of bedrock erosion under glaciers, we
    infer a distribution of potential glacier erosion in the Alps over the last
    glacial cycle from 120\,000 years ago to the present.
    %
    Depsite large uncertainties related to the climate history of the Alps and
    glacier erosion processes, the resulting modelled patterns of glacier
    erosion shows persistent features (hopefully). The cumulative imprint of
    the last glacial cycle shows peak erosion at the mouth of the large Alpine
    valleys where glaciers are modelled to have flown with the highest
    velocity. The modelled erosion pattern varies significantly through the
    glacial cycle, but surprisingly the total rate of glacier erosion is
    modelled to be relatively stable during the entire glacial cycle.  While
    glacial maxima lead to high erosion rates at low elevations, the
    high-elevation areas are typically preserved under cold-based ice during
    these periods, but are modelled to have experience mode intense erosion
    during periods of intermediate glaciation extent.
    %
    This result indicate that different landscapes of the same mountain range
    most likely correspond to different time periods, and explains the
    diversity of glacial landscapes in the Alps.

>>>>

\title{Last glacial cycle glacier erosion potential in the Alps}
\author[1]{Julien Seguinot}
\author[2]{Ian Delaney}
\affil[1]{Anafi, Greece} % {\small (juseg \emph{at} posteo.eu)}}
\affil[2]{Institute of Earth Surface Dynamics, University of Lausanne, Switzerland}

@@

    The glacial landscape of the Alps has fascinated generations of explorers,
    artists, mountaineers and scientists with its diversity, including
    erosional features of all scales from high-mountain cirques, to steep
    glacial valleys and large over-deepened basins. Using previous glacier
    modelling results, and empirical inferences bedrock erosion under modern
    glaciers, we compute a distribution of potential glacier erosion in the Alps
    over the last glacial cycle from 120\,000 years ago to the present.
    %
    Despite large uncertainties related to the climate history of the Alps and
    unconstrained glacier erosion processes, the resulting modelled patterns of
    glacier erosion shows persistent features. The cumulative imprint of
    the last glacial cycle shows a very strong localization of glacier erosion
    with local maxima at the mouths of major Alpine valleys and some other
    upstream sections where glaciers are modelled to have flown with the
    highest velocity. The modelled erosion rates vary significantly through the
    glacial cycle, but show paradoxically little relation to the glaciation
    extent. Phases of glacier advance and maximum extension, characterized by
    low air and glacier temperatures, see a localization of rapid erosion rates
    at low elevation, while intra-montane areas are modelled to date from phases
    of intermediate glacier cover. The modelled erosion rates peak during
    deglaciation phases, when higher air and glacier temperature enhance
    glacier dynamics.
    %
    Our results depict the Alpine glacier erosion landscape as a
    time-transgressive patchwork, with different erosional landscapes of the
    same mountain range corresponding to different glaciation stages and time
    periods, thus offering insight into its diversity.
